---
title: "Tidy Tuesday - Week 35"
subtitle : "Lemurs"
author: "Christine Hedde - von Westernhagen"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: paper
    highlight: pygment
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",
                      fig.pos = "H")

library(tidyverse)
library(here)
library(tidygraph)
library(ggraph)
library(ggtext)

extrafont::loadfonts(device = "win", quiet = TRUE)

data_raw <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-24/lemur_data.csv')

```

# Goal

Network plot of hybrid species relations: Which species are connected to one another through an offspring?

# Data Exploration

```{r data_expl}

# How many hybrids are there?
data_raw %>% 
  distinct(dlc_id, hybrid) %>% 
  count(hybrid)

# How many parent species are there?
data_raw %>% 
  distinct(dam_taxon) %>% 
  nrow()

data_raw %>% 
  distinct(sire_taxon) %>%
  nrow()

# What are the taxon combinations of the parents?
data_raw %>%
  distinct(dlc_id, hybrid, dam_taxon, sire_taxon) %>% 
  filter(hybrid=="Sp") %>% 
  count(dam_taxon, sire_taxon) %>% 
  arrange(desc(n))

```

To conclude: 173 of 2270 individuals are hybrids of two species. For their 27 female and 26 male parent-species we can observe 28 pairwise combinations. However, the male parent is often unkown, "pater semper incertus est". Most of the parents are hybrids themselves (EUL).

# Make it a network object

```{r make_net}

# make an edgelist and turn into tidygraph object
net <- data_raw %>% 
  distinct(dlc_id, hybrid, dam_taxon, sire_taxon) %>% 
  filter(hybrid=="Sp") %>% 
  count(dam_taxon, sire_taxon) %>% 
  arrange(desc(n)) %>% 
  rename(from = dam_taxon, to = sire_taxon) %>% 
  as_tbl_graph(directed = FALSE)

# load translation table of taxonomic codes
taxon_codes <- 
  read_delim("taxon_codes.csv", "|", trim_ws = TRUE) %>% 
  janitor::clean_names() %>% 
  select(2:4) %>% 
  tail(-1)

# edits to the final data set
net_final <- net %>% 
  activate(nodes) %>% 
  rename(taxon = name) %>% 
  left_join(taxon_codes) %>% 
  mutate(common_name = case_when(taxon == "EUL" ~ "Eulemur hybrid",
                                 taxon == "VAR" ~ "Varecia hybrid",
                                 is.na(common_name) ~ "Unknown",
                                 TRUE ~ common_name) %>% 
           str_wrap(., 14),
         family = case_when(taxon %in% c("CMEAD", "MMUR", "MZAZ") ~ "Cheirogaleidae",
                            taxon == "DMAD" ~ "Daubentoniidae",
                            str_detect(taxon, "^E|^H|^LC|^V") ~ "Lemuridae",
                            taxon == "GMOH" ~ "Galago",
                            taxon == "LTAR" ~ "Loris",
                            taxon %in% c("NCOU", "NPYG") ~ "Nycticebus",
                            taxon == "OGG" ~ "Otolemur",
                            taxon == "PCOQ" ~ "Indriidae",
                            taxon == "PPOT" ~ "Perodicticus",
                            TRUE ~ "Unknown"),
         species = ifelse(family != "Unknown", 
                          str_extract_all(latin_name, "^[A-Z][a-z]+"),
                          "Unknown"),
         degree = centrality_degree(weights = n, loops = TRUE))

net_final %>% 
  activate(nodes) %>% 
  as_tibble()

net_final %>% 
  activate(edges) %>% 
  as_tibble()

```

# Plot

```{r plot, fig.width=10, fig.height=10.5, fig.cap="Network of hybrid lemurs' parental species", fig.alt=fig.alt}

fig.alt <- "The Duke Lemur Center has hosted more than 2000 Lemurs across 27 species since 1968. Around eight percent of the animals are hybrids of different species. The graph shows the taxonomic relations between their parents, who are often hybrids themselves, as a circular network consisting of 13 nodes. Most connections run between hybrid species. However, in many cases the father's species could not be identified, 'pater semper incertus est'."


pal <- c("#FDF8E5", "#EBA843", "#ff6b13", "#7D5F45", "#42474B", "#B2C438", "#0B5587")


ggraph(net_final, layout = "linear", sort.by = degree, circular = TRUE) +
  geom_edge_arc(aes(width = stat(index)-(stat(index))^2, color = n, alpha = n+0.5),
                lineend = "round",
                start_cap = circle(3, "mm"),
                end_cap = circle(3, "mm")) +
  geom_edge_loop(aes(width = sqrt(stat(index))-stat(index), color = n, alpha = n+0.5),
                 lineend = "round",
                 start_cap = circle(3, "mm"),
                 end_cap = circle(3, "mm")) +
  geom_node_text(aes(label = common_name),
                 size = 4, 
                 color = pal[1],
                 family = "Impact") +
  scale_edge_color_gradientn(colors = rev(pal[1:3])) +
  scale_edge_alpha(guide = "none") +
  scale_edge_width(range = c(.7, 2.5), guide = "none") +
  guides(edge_color = guide_edge_colorbar(title = "Number of offspring", 
                                          title.position = "top",
                                          barwidth = 7,
                                          barheight = .5)) +
  labs(title = "Lemur Hybrids", 
       subtitle = "The Duke Lemur Center has hosted more than 2000 Lemurs across 27 species since 1968. Around eight percent of the animals are hybrids of different species. The graph shows the taxonomic relations between their parents, who are often hybrids themselves. However, in many cases the father's species could not be identified, 'pater semper incertus est'.",
       caption = "Data source: The Duke Lemur Center, provided by Jesse Mostipak\nPlot: @christinehvw") +
  expand_limits(x = c(-1.1, 1.1)) +
  theme_graph() +
  theme(plot.background = element_rect(fill = pal[5]),
        text = element_text(size = 11, color = pal[1], family = "Arial Nova"),
        plot.margin = unit(c(0,6,2,2), "mm"),
        legend.position = c(.9,0.04), 
        legend.direction = "horizontal",
        legend.title.align = 0.5,
        plot.title = element_textbox_simple(size = 28, 
                                            family = "Monoton", 
                                            color = pal[1],
                                            margin = margin(4,4,4,1, "mm")),
        plot.subtitle = element_textbox_simple(size = 10, 
                                               family = "Arial Nova", 
                                               color = pal[1],
                                               margin = margin(0,120,5,1, "mm")),
        plot.caption = element_text(family = "Arial Nova",
                                    face = "italic",
                                    color = "black",
                                    hjust = 0,
                                    margin(0,0,4,1, "mm"))) 

```


```{r save, eval=FALSE, include=FALSE}

ggsave(filename = "plot.png", dpi = 400,
       width = unit(10, "in"), height = unit(10.5, "in"))

```


